---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# swfscAirDAS

<!-- badges: start -->
[![Travis build status](https://travis-ci.org/smwoodman/swfscAirDAS.svg?branch=master)](https://travis-ci.org/smwoodman/swfscAirDAS)
[![AppVeyor build status](https://ci.appveyor.com/api/projects/status/github/smwoodman/swfscAirDAS?branch=master&svg=true)](https://ci.appveyor.com/project/smwoodman/swfscAirDAS)
<!-- badges: end -->

This package contains functions for processing and analyzing aerial survey DAS data (AirDAS) generated by turtle4d (and related programs) at the Southwest Fisheries Science Center.

## Installation

You can install swfscAirDAS from [GitHub](https://github.com/smwoodman/swfscAirDAS) with:

``` r
# install.packages("devtools")
devtools::install_github("smwoodman/swfscAirDAS")
```

## Usage

```{r example}
library(swfscAirDAS)
# Get file paths of sample files included in the package
y <- system.file("airdas_sample.das", package = "swfscAirDAS")
y.eff.randpicks <- system.file("airdas_sample_randpicks.csv", package = "swfscAirDAS")

# Read and process AirDAS file, i.e. read AirDAS data into a data frame and add info columns
y.read <- airdas_read(y)
y.proc <- airdas_process(y.read)

# Alternatively, the file path can be passed directly to airdas_process
y.proc <- airdas_process(y) 

# Summarize sighting information
y.sight <- airdas_sight(y.proc)

# Chop continuous effort sections into effort (modeling) segments, 
#   and summarize information for each segment
y.eff <- airdas_effort(y.proc, 3, sp.codes = c("mn", "bm", "qq"), randpicks.load = y.eff.randpicks)
y.eff.segdata <- y.eff[[1]]
y.eff.siteinfo <- y.eff[[2]]
y.eff.randpicks <- y.eff[[3]]
```

To read in and concatenate mutliple AirDAS files, currently you can use the following code:

```{r, eval = FALSE}
y.proc <- do.call(rbind, lapply(list.files("filepath/to/airdas_file_folder", full.names = TRUE), airdas_read))
```

## Vignettes

```r
# To see the list of available vignettes
browseVignettes("swfscAirDAS") 

# To open a specific vignette
vignette("AirDAS_format")
```

* AirDAS data format requirements ([AirDAS_format](https://github.com/smwoodman/swfscAirDAS/blob/master/vignettes/AirDAS_format.Rmd))
