---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# swfscAirDAS

<!-- badges: start -->
[![Travis build status](https://travis-ci.com/smwoodman/swfscAirDAS.svg?branch=master)](https://travis-ci.com/smwoodman/swfscAirDAS)
[![AppVeyor build status](https://ci.appveyor.com/api/projects/status/github/smwoodman/swfscAirDAS?branch=master&svg=true)](https://ci.appveyor.com/project/smwoodman/swfscAirDAS)
<!-- badges: end -->

Warning: swfscAirDAS is still under development! Please consult with the author (sam.woodman@noaa.gov) before using it.

This package contains functions designed for processing and analyzing aerial survey DAS data (AirDAS) generated by the Southwest Fisheries Science Center using one of the following programs: PHOCOENA, SURVEY, CARETTA, or TURTLE (such as TURTLEP or TURTLE4D). Functionality includes reading AirDAS data into a data frame, processing this data (extracting state and condition information for each AirDAS event), and summarizing sighting and effort information. Plans for future development include: generating tables summarizing the information in the AirDAS file, plotting AirDAS data, and generating files needed for doing density estimation.

## Installation

You can install `swfscAirDAS` from [GitHub](https://github.com) with:

``` r
# install.packages("devtools")
devtools::install_github("smwoodman/swfscAirDAS")
```

## AirDAS format

You can [download the PDF](https://github.com/smwoodman/swfscAirDAS/blob/master/inst/AirDAS_Format.pdf) describing the AirDAS data format requirements of `swfscAirDAS`.

## Usage

First, you must read and process the AirDAS data

```{r example}
# library(swfscAirDAS)
# # Get file paths of sample files included in the package
# y <- system.file("airdas_sample.das", package = "swfscAirDAS")
# y.eff.randpicks <- system.file("airdas_sample_randpicks.csv", package = "swfscAirDAS")
# 
# # Read and process AirDAS file, i.e. read AirDAS data into a data frame and add info columns
# y.read <- airdas_read(y)
# y.proc <- airdas_process(y.read)
# 
# # Alternatively, the file path can be passed directly to airdas_process
# y.proc <- airdas_process(y)
```

Note that `airdas_read` can read multiple files simultaneously

```{r}
# y.read.mult <- airdas_read(c(y, y))
```

After processing, you can summarize the processed AirDAS data. There are two available methods for chopping and summarizing effort: "equallength" and "condition". The method "equallength" chops continuous effort sections into equal length effort (modeling) segments, while the method "condition" chops continuous effort sections in segments every time a condition changes. 

```{r}
# # Summarize sighting information
# y.sight <- airdas_sight(y.proc)
# 
# # Chop and summarize effort using "equallength" method
# y.eff <- airdas_effort(
#   y.proc, method = "equallength", sp.codes = c("mn", "bm"), 
#   seg.km = 3, randpicks.load = y.eff.randpicks
# )
# y.eff.segdata <- y.eff[[1]]
# y.eff.siteinfo <- y.eff[[2]]
# y.eff.randpicks <- y.eff[[3]]
# 
# # Chop and summarize effort using "condition" method
# y.eff.cond <- airdas_effort(
#   y.proc, method = "condition", sp.codes = c("mn", "bm"), 
#   seg.km.min = 0.05
# )
```

You can also check that your AirDAS data has accepted formatting and values:

```{r, eval=FALSE}
# airdas_check(y)
```

